Binary companions of red giant stars in \apogee\ \DR: \\ Search method and catalog of $>$4,500 systems

### 2018-03-19 paper/1-catalog/twoface-catalog.tex
Placeholder for upper RGB
TODOs
new figures and shite
new figure
logg teff

% define macros for text
\newcommand{\apogee}{\project{\acronym{APOGEE}}}
\newcommand{\sdssiv}{\project{\acronym{SDSS-IV}}}
\newcommand{\thejoker}{\project{The~Joker}}
\newcommand{\thecannon}{\project{The~Cannon}}
\newcommand{\DR}{\acronym{DR14}}
\newcommand{\DRtw}{\acronym{DR12}}
\newcommand{\RC}{\acronym{RC}}
\newcommand{\RGB}{\acronym{RGB}}
\newcommand{\quantile}[2]{\ensuremath{\textrm{Q}_{#1}}\left(#2\right)}
\newcommand{\pdf}{\textit{pdf}}

@@

Data taken as part of the \apogee\ survey are reduced with a multi-step data
reduction pipeline that ultimately solves for the stellar parameters, chemical
abundances, and radial velocities for each target
(\citealt{Nidever:2015}).
Most relevant for this work, the visit radial velocities (RVs) are determined
using an iterative scheme: the individual visit spectra are combined using
initial guesses for the relative RVs into a coadded spectrum, which is then used
to re-derive the relative visit velocities.
The stellar parameters---surface gravity, $\log g$, and effective temperature,
$T_{\textrm{eff}}$---and the chemical abundances are determined from the coadded
spectrum as a part of the \apogee\ Stellar Parameters and Chemical Abundances
Pipeline (\acronym{ASPCAP}; \citealt{ASPCAP}).

@@

We use the primary data products from \apogee\ \DR\ (i.e. the \texttt{allStar}
and \texttt{allVisit} files) which contain 258,475 unique source IDs
(\texttt{APOGEE\_ID}) and 1,054,381 unique visits.
We select all stars with $\geq 3$ visits that each pass a set of quality cuts,
described below.
For each visit, we require that the visit velocity uncertainty is $< 100~\kms$
(\texttt{VRELERR}) and the following bits are not set in the \texttt{STARFLAGS}
bitmask: \texttt{PERSIST\_HIGH}, \texttt{PERSIST\_JUMP\_POS},
\texttt{PERSIST\_JUMP\_NEG}, \texttt{VERY\_BRIGHT\_NEIGHBOR}, \texttt{LOW\_SNR}.
For each star, we require that $0 < \log g < 4$ and the following bits are not
set in the \texttt{ASPCAPFLAGS} bitmask: \texttt{STAR\_BAD}.
After these cuts, and the requirement of $\geq 3$ visits for a given star, we
are left with \nvisits\ visits for \nstars\ unique sources.
\figurename~\ref{fig:nvisits} shows the number of stars in several bins of
number of visits that pass the above quality cuts: 92\% of the stars in
our parent sample have $< 8$ visits.
\figurename~\ref{fig:loggteff} shows the stellar parameters of all \nstars\
stars in the parent sample used in this work.

% Notebook: "Numbers of visits and sample overview"
\begin{figure}[h]
\begin{center}
\includegraphics[width=\textwidth]{logg_teff_feh.pdf}
\end{center}
\caption{%
\textit{Left panel}: Number of stars in bins of iron abundance,
$[\textrm{Fe}/\textrm{H}]$, that pass the quality cuts described in
\sectionname~\ref{sec:data}.
\textit{Right panel}: Distribution of stars in our sample in stellar parameters
log-surface-gravity, $\log g$, and effective temperature, $T_{\textrm{eff}}$,
with points colored by the iron abundance.
\label{fig:loggteff}
}
\end{figure}

@@

We use 1,825 stars with $>10$ visits and $\log g < 2$ (to avoid upper RGB stars
that have large intrinsic jitter) and maximize the above likelihood to determine
better hyperparameters for the log-excess-variance parameter distribution.
\figurename~\ref{fig:infer-jitter} shows the distribution corresponding to the
maximum-likelihood hyperparameters, $\alpha^* = (\mu_y^*, \sigma_y^*) = (9.50,
1.64)$.
These values are consistent with the estimated systematic floor of the visit
velocity uncertainties of $\approx 100$--$200~\mps$ estimated from stars
observed multiple times and on multiple plates (\citealt{Nidever:2015}).

@@

Here we highlight a few interesting results from visualizing the properties of
companions in the unimodal and bimodal samples defined above (see \sectionname
s~\ref{sec:catalog-unimodal} and \ref{sec:catalog-multimodal}).
In all figures, we only plot companions with inferred orbits that pass our
visual inspection (\texttt{clean\_flag == 0}).
We additionally restrict to stars with $\log g > 2$ to avoid upper RGB stars
that may have coherent oscillations that mimic RV modulations from companions.

@@

% Notebook: "Results"
\begin{figure}[h]
\begin{center}
\includegraphics[width=\textwidth]{mass}
\end{center}
\caption{%
\textit{Left panel:} Estimated minimum companion mass and primary mass for a
subset of the systems in the high-$K$, unimodal converged sample (31 systems;
black circles), unimodal unconverged sample (38 systems; grey circles), and
bimodal sample (25 systems; blue squares and lines) with $\log g > 2$ and
previously measured primary masses (\citealt{Ness:2015}).
Lines connecting markers (blue) show estimates from each period mode for each
system with bimodal period samplings.
Upper, dashed line shows the equal-mass curve, and lower dashed line shows the
Hydrogen-burning limit.
\textit{Right panel:} Ratio of primary stellar radius, $R_1$, to projected
separation of the two bodies, $a\,\sin i$, and minimum mass ratio,
$q_{\textrm{min}}$, for the same systems and markers in left panel.
\label{fig:mass}
}
\end{figure}

@@

@@

One of these studies focused on decomposing spectra of MS stars into mixtures of
stellar spectra (\citealt{El-Badry:2018}).
Conceptually, this method works because (a) for two unequal-mass stars,
unexpected absorption lines will appear superimposed on the brighter star's
spectrum, and (b) for close to equal-mass stars, the line depths and ratios will
not be well-matched by a single stellar model.
Using this technique, they identified thousands of candidate MS binaries and
trinaries, but did not consider giant stars ($\log g > 4$).
This sample is therefore complimentary to and non-overlapping with the catalog
presented in this work.

>>>>

% define macros for text
\newcommand{\apogee}{\project{\acronym{APOGEE}}}
\newcommand{\sdssiv}{\project{\acronym{SDSS-IV}}}
\newcommand{\thejoker}{\project{The~Joker}}
\newcommand{\thecannon}{\project{The~Cannon}}
\newcommand{\DR}{\acronym{DR14}}
\newcommand{\DRtw}{\acronym{DR12}}
\newcommand{\RC}{\acronym{RC}}
\newcommand{\RGB}{\acronym{RGB}}
\newcommand{\quantile}[2]{\ensuremath{\textrm{Q}_{#1}}\left(#2\right)}
\newcommand{\pdf}{\textit{pdf}}
\newcommand{\logg}{\ensuremath{\log g}}
\newcommand{\Teff}{\ensuremath{T_{\textrm{eff}}}}

@@

Data taken as part of the \apogee\ survey are reduced with a multi-step data
reduction pipeline that ultimately solves for the stellar parameters, chemical
abundances, and radial velocities for each target
(\citealt{Nidever:2015}).
Most relevant for this work, the visit radial velocities (RVs) are determined
using an iterative scheme: the individual visit spectra are combined using
initial guesses for the relative RVs into a coadded spectrum, which is then used
to re-derive the relative visit velocities.
The stellar parameters---surface gravity, \logg, and effective temperature,
\Teff---and the chemical abundances are determined from the coadded
spectrum as a part of the \apogee\ Stellar Parameters and Chemical Abundances
Pipeline (\acronym{ASPCAP}; \citealt{ASPCAP}).

@@

We use the primary data products from \apogee\ \DR\ (i.e. the \texttt{allStar}
and \texttt{allVisit} files) which contain 258,475 unique source IDs
(\texttt{APOGEE\_ID}) and 1,054,381 unique visits.
We select all stars with $\geq 3$ visits that each pass a set of quality cuts,
described below.
For each visit, we require that the visit velocity uncertainty is $< 100~\kms$
(\texttt{VRELERR}) and the following bits are not set in the \texttt{STARFLAGS}
bitmask: \texttt{PERSIST\_HIGH}, \texttt{PERSIST\_JUMP\_POS},
\texttt{PERSIST\_JUMP\_NEG}, \texttt{VERY\_BRIGHT\_NEIGHBOR}, \texttt{LOW\_SNR}.
For each star, we require that $0 < \logg < 4$ and the following bits are not
set in the \texttt{ASPCAPFLAGS} bitmask: \texttt{STAR\_BAD}.
After these cuts, and the requirement of $\geq 3$ visits for a given star, we
are left with \nvisits\ visits for \nstars\ unique sources.
\figurename~\ref{fig:nvisits} shows the number of stars in several bins of
number of visits that pass the above quality cuts: 92\% of the stars in
our parent sample have $< 8$ visits.
\figurename~\ref{fig:loggteff} shows the stellar parameters of all \nstars\
stars in the parent sample used in this work.

% Notebook: "Numbers of visits and sample overview"
\begin{figure}[h]
\begin{center}
\includegraphics[width=\textwidth]{logg_teff_feh.pdf}
\end{center}
\caption{%
\textit{Left panel}: Number of stars in bins of iron abundance,
$[\textrm{Fe}/\textrm{H}]$, that pass the quality cuts described in
\sectionname~\ref{sec:data}.
\textit{Right panel}: Distribution of stars in our sample in stellar parameters
log-surface-gravity, \logg, and effective temperature, \Teff, with points
colored by the iron abundance.
\label{fig:loggteff}
}
\end{figure}

@@

We use 1,825 stars with $>10$ visits and $\logg > 2$ (to avoid upper RGB stars
that have large intrinsic jitter) and maximize the above likelihood to determine
better hyperparameters for the log-excess-variance parameter distribution.
\figurename~\ref{fig:infer-jitter} shows the distribution corresponding to the
maximum-likelihood hyperparameters, $\alpha^* = (\mu_y^*, \sigma_y^*) = (9.50,
1.64)$.
These values are consistent with the estimated systematic floor of the visit
velocity uncertainties of $\approx 100$--$200~\mps$ estimated from stars
observed multiple times and on multiple plates (\citealt{Nidever:2015}).

@@

Here we highlight a few interesting results from visualizing the properties of
companions in the unimodal and bimodal samples defined above (see \sectionname
s~\ref{sec:catalog-unimodal} and \ref{sec:catalog-multimodal}).
In all figures, we only plot companions with inferred orbits that pass our
visual inspection (\texttt{clean\_flag == 0}).
We additionally restrict to stars with $\logg > 2$ to avoid upper RGB stars
that may have coherent oscillations that mimic RV modulations from companions.

\todo{here:} \figurename~\ref{fig:logg-teff-p} ...

% Notebook: "Results"
\begin{figure}[h]
\begin{center}
\includegraphics[width=\textwidth]{mass}
\end{center}
\caption{%
\textit{Left panel:}
\textit{Right panel:}
\label{fig:logg-teff-p}
}
\end{figure}

@@

% Notebook: "Results"
\begin{figure}[h]
\begin{center}
\includegraphics[width=\textwidth]{mass}
\end{center}
\caption{%
\textit{Left panel:} Estimated minimum companion mass and primary mass for a
subset of the systems in the high-$K$, unimodal converged sample (31 systems;
black circles), unimodal unconverged sample (38 systems; grey circles), and
bimodal sample (25 systems; blue squares and lines) with $\logg > 2$ and
previously measured primary masses (\citealt{Ness:2015}).
Lines connecting markers (blue) show estimates from each period mode for each
system with bimodal period samplings.
Upper, dashed line shows the equal-mass curve, and lower dashed line shows the
Hydrogen-burning limit.
\textit{Right panel:} Ratio of primary stellar radius, $R_1$, to projected
separation of the two bodies, $a\,\sin i$, and minimum mass ratio,
$q_{\textrm{min}}$, for the same systems and markers in left panel.
\label{fig:mass}
}
\end{figure}

\subsection{Coherent...}

\todo{Coherent oscillations?}

@@

\todo{Discuss tidal circ and mention worth pursuing. Don't say we are going to
do this. Emphasize that the data contain information about those things.}

\todo{High-mass companions}

@@

One of these studies focused on decomposing spectra of MS stars into mixtures of
stellar spectra (\citealt{El-Badry:2018}).
Conceptually, this method works because (a) for two unequal-mass stars,
unexpected absorption lines will appear superimposed on the brighter star's
spectrum, and (b) for close to equal-mass stars, the line depths and ratios will
not be well-matched by a single stellar model.
Using this technique, they identified thousands of candidate MS binaries and
trinaries, but did not consider giant stars ($\logg > 4$).
This sample is therefore complimentary to and non-overlapping with the catalog
presented in this work.
